Investigating Determinism of Gaming Engines for Autonomous Vehicle Verification

### 2019-12-23 paper/3_Methodology.tex

% \noindent 
In this section a method for determining the level of \textit{simulation variance} is presented along with experimental parameters that may influence it, see Fig.~\ref{method_diagram}. The method includes scenarios that exercise different actor types and functional callbacks of the game engine that may influence thread scheduling. To replicate high computational loads, and hence game loop update rates, that may be experienced in high realism simulations, resource utilisation software is used at runtime for both central and graphical processing units.

@@

Increasing the physics time-step calculations, which is increasing the number of calculations per unit time, would improve the level of simulation variance since the physics sequence will be more finely defined; but on the other hand this would be more computationally expensive. However, in such applications of V\&V of CAV simulations, the computational cost should be of less concern compared to the importance of repeatability of tests.

If the experiment setup includes the usage of the physics engine's AI then altering the navigation mesh settings in the engine, like increasing the granularity of the mesh or changing the mesh type; can improve the level of simulation variance as a result of allowing the AI to navigate in a more well defined space.

@@

Running in editor or deployed mode can sometimes cause massive differences in performance, with deployed mode being worse, mostly due to the way settings get packaged when in deployed mode.

>>>>

% \noindent 
In this section a method for determining the \textit{simulation variance} is presented along with experimental parameters that may influence it, see Fig.~\ref{method_diagram}. The method includes scenarios that exercise different actor types and functional callbacks of the game engine that may influence thread scheduling. To replicate high computational loads, and hence game loop update rates, that may be experienced in high realism simulations, resource utilisation software is used at runtime for both central and graphical processing units.

@@

Increasing the physics time-step calculations, which is increasing the number of calculations per unit time, would improve the simulation variance since the physics sequence will be more finely defined; but on the other hand this would be more computationally expensive. However, in such applications of V\&V of CAV simulations, the computational cost should be of less concern compared to the importance of repeatability of tests.

If the experiment setup includes the usage of the physics engine's AI then altering the navigation mesh settings in the engine, like increasing the granularity of the mesh or changing the mesh type; can improve the simulation variance as a result of allowing the AI to navigate in a more well defined space.

@@

Running in editor or deployed mode can sometimes cause differences in performance, with deployed mode being worse, mostly due to the way settings get packaged when in deployed mode.

### 2019-12-23 paper/4_CaseStudy.tex

\subsection{Problem And Hypothesis}
We want to see whether using Unreal Engine 4 (UE4) along with CARLA \cite{CARLA_paper} for V\&V of CAV simulations is deterministic or not; if not then what is the level of simulation variance that we will operate with, is it acceptable or not, and whether it worsens with stressing GPU and CPU.

Our hypothesis is that since UE4 is a gaming engine it is not deterministic, however, the level of simulation variance would be very low to acceptable limits that allows verification engineers to use it confidently and assume that tests would be repeatable.
This behaviour is expected not to hold as the level of computing stress increases.

@@

Note the dotted red line in the plot shows the millimeter (mm) level. 
Below this level we assume that the level of simulation variance is low enough that it would not affect our assertion checking for V\&V of CAV simulations purposes, that it is sufficient to assume the tests are deterministic and repeatable. 
This line can be moved up or down and it is entirely up to verification engineers where they want to define the level below which tests' variance is negligibale and it is suffcient to assume that tests are deterministic.

@@

Next, given the pedestrians were showing a very robust and low level of simulation variance, we ran tests for pedestrians colliding (i.e. with paths intersection) for different look ahead distances to exploit whether that would cause any changes, but yet as it can be seen by lines Test IDs' 6,7 and 8 that their performance is consistent for all of the different tests.

>>>>

\subsection{Problem And Hypothesis}
We want to see whether using Unreal Engine 4 (UE4) along with CARLA \cite{CARLA_paper} for V\&V of CAV simulations is deterministic or not; if not then what is the simulation variance that we will operate with, is it acceptable or not, and whether it worsens with stressing GPU and CPU.

Our hypothesis is that since UE4 is a gaming engine it is not deterministic, however, the simulation variance would be very low to acceptable limits that allows verification engineers to use it confidently and assume that tests would be repeatable.
This behaviour is expected not to hold as the level of computing stress increases.

@@

Note the dotted red line in the plot shows the millimeter (mm) level. 
Below this level we assume that the simulation variance is low enough that it would not affect our assertion checking for V\&V of CAV simulations purposes, that it is sufficient to assume the tests are deterministic and repeatable. 
This line can be moved up or down and it is entirely up to verification engineers where they want to define the level below which tests' variance is negligibale and it is suffcient to assume that tests are deterministic.

@@

Next, given the pedestrians were showing low simulation variance, we ran tests for pedestrians colliding (i.e. with paths intersection) for different look ahead distances to exploit whether that would cause any changes, but yet as it can be seen by lines Test IDs' 6,7 and 8 that their performance is consistent for all of the different tests.
